### 2023-04-17 chapters/depsynth/conclusion.tex

\section{Conclusion}\label{s:conclusion}

\depsynth offers a new programming model for building crash-consistent storage systems.
By offering a high-level angelic programming model for crash consistency,
and automatically synthesizing low-level dependency rules to realize that model,
\depsynth lowers the burden of building reliable storage systems.
We believe that this work presents a promising direction for building systems software
with the aid of automatic programming tools
to resolve challenging nondeterminism and persistence problems.