### 2023-04-17 chapters/depsynth/impl.tex

\section{Implementation}\label{s:impl}

We implement both the \depsynth algorithm
and the storage systems we study in \cref{sec:eval}
in Rosette~\cite{torlak:rosette},
an extension of Racket~\cite{felleisen:racket}
with support for verification and synthesis.
Using Rosette as our host language gives us symbolic evaluation of the storage system implementation for free,
and simplifies implementing the \crashconsistentalg query in \cref{fig:alg:depsynth}.
The choice of Rosette and Racket is not fundamental;
recent work has shown how to extend the symbolic evaluation approach to languages
such as Python~\cite{sigurbjarnarson:yggdrasil} or C~\cite{nelson:serval}
in which storage systems are more commonly implemented.

\paragraph{Ordering.}
The \depsynthalg algorithm in \cref{fig:alg:depsynth}
is sensitive to the order in which \textsc{NextTest} chooses tests to generate dependency rules for.
Our implementation chooses tests in increasing order of size,
minimizing the number of happens-before graphs for \rulesfortest to explore.
Similarly, \rulesfortest is sensitive to the order it considers writes (\phaseone)
and edges (\phasetwo).
In both cases, we exploit the following observation: 
while an execution that persists writes in program order is not \emph{required} to be crash consistent
(e.g., because storage systems might selectively buffer or coalesce writes),
it is often so in practice.
\rulesfortest therefore prefers to choose writes in \phaseone in program order,
and prefers to remove edges in \phasetwo that contradict program order.

\paragraph{Reducing solver queries.}
Both \phaseone and \phasetwo in \cref{fig:alg:rulesfortests}
have symmetry in their search space:
for a fixed pair of writes $w_1$ and $w_2$,
there are many different branches of \phaseone
that try to order $w_2$ after $w_1$,
and many different branches of \phasetwo
that try to remove the edge $(w_1, w_2)$ from a happens-before graph.
If we can determine ahead of time that such a choice for those writes is always doomed to fail,
we can avoid considering these choices at all
and so save the cost of an SMT solver query by \crashconsistentalg.
Our implementation of \rulesfortest
uses an SMT solver to
pre-compute a set of \emph{necessary} ordering edges---%
edges which \emph{must} be in the happens-before graph---%
and uses that set to short-circuit \crashconsistentalg.

\if 0

%\paragraph{Avoiding \multisearch}
%In \autoref{alg:top-level}, we described our top-level algorithm as one that
%first uses an efficient per-litmus-test search before defaulting to a
%slower search for dependency graphs over multiple litmus tests.
%However, even in the presence of rule conflicts, \depsynth can still avoid
%performing the slower search in most cases. The way \depsynth achieves this
%is by searching for new graphs for individual conflicting litmus tests
%that, when extrapolated into rules, do not cause a conflict.
%
%Specifically, if \depsynth encounters a set of litmus tests $\{L_0, L_1, ...\}$
%whose generated rules conflict, \depsynth first tries to generate a new graph and rules
%for $L_0$ such that these new rules jj

\paragraph{Rosette}
\todo{Talk about how rosette is used in \depsynth}

\paragraph{Limiting Queries with Preprocessing}

% TODO explain necessary paths
At each state of the dependency graph search in \sccsearch, \depsynth
performs an SMT query to determine if the current candidate graph
is sufficient. These queries save time by allowing \sccsearch to avoid
searching unnecessary spaces, but the queries themselves are not free.
Though the queries are fairly lightweight, as the free variables are on
the order of 10s of boolean values, they still introduce noticeable overhead.

Generally, the results of sufficiency queries for similar graphs are similar:
one example is that, as previously discussed, subgraphs of insufficient graphs
are always insufficient.
In fact, we noticed that in most cases, certain paths between nodes are
\textit{necessary} in the final output of \sccsearch, meaning that
removing all paths between these nodes from a graph would immediately
cause it to become insufficient.
Intuitively, these necessary paths exist whenever all solution graphs
to our search must contain such a path.
If \depsynth can precompute the set of necessary paths,
we can avoid querying for sufficiency over graphs that lack any.

Formally, we say there is a \textit{necessary path} from $w_0$ to $w_1$, whenever
there \textit{does not} exist a schedule with $\syncbool_0\mapsto\true$
and $\syncbool_1\mapsto\false$ that satisfies the crash consistency property
(i.e. committing $w_0$ before $w_1$ is never crash consistent).
If such a schedule does not exist, it means that for every schedule satisfying
the property, $\syncbool_0\implies\syncbool_1$, and so any solution graph to our
search must contain a path from $w_0$ to $w_1$.
When a graph lacks any path from $w_0$ to $w_1$, it can be pruned
without checking for sufficiency.
For each pair of writes $w_i,w_j$,
we can determine if there is a necessary path from $w_i$ to $w_j$
by querying over all schedules with
$\syncbool_j\mapsto\true$ and $\syncbool_j\mapsto\false$.

% TODO explain how nec paths reduce the number of queries we have to make
Though this precomputation would prevent some sufficiency queries,
it requires checking for necessary paths between all nodes.
Naively, this would take $|\omega|(|\omega|-1)$ queries for set of writes
$\omega$, but we can avoid most of these queries by taking advantage of the fact
that our SMT solver returns a concrete schedule whenever a necessary
path \textit{does not} exist.
Specifically, when query for writes $w_0$ and $w_1$ determines that there is
a schedule that satisfies property $\mathcal{P}$ where $\syncbool_0\mapsto\true$
and $\syncbool_1\mapsto\false$, consider all other pairs $i,j$ in the schedule
where $\syncbool_i\mapsto\true$ and $\syncbool_j\mapsto\false$.
From this schedule, we can also deduce that there is no necessary path
from $w_i$ to $w_j$. Using this optimization, we can compute all necessary paths
on the order of $O(|\omega|)$ queries.

An added benefit of necessary edges is that \sccsearch can start from 
a smaller graph than the complete graph. Specifically, if there exists
a necessary path from $w_0$ to $w_1$, then we can remove edge $(w_1, w_0)$
from the starting graph, as any sufficient graph with both a path from
$w_0$ to $w_1$ and edge $(w_1, w_0)$ must contain a cycle.
In some cases, this means that not all starting nodes share a single SCC.

\paragraph{Prioritizing choices in \sccsearch}
\todo{These few paragraphs could go in the technical section, just not sure}
In \autoref{alg:scc-search}, we describe an algorithm that, at each step,
chooses an arbitrary node to order with respect to other nodes in its
strongly connected component.
In practice, \depsynth prioritizes ordering nodes in two ways:
\begin{enumerate}
  \item Order nodes as they would be ordered with previously generated rules.
  \item Order nodes from newest to oldest.
\end{enumerate}

Since \depsynth keeps the set of rules it has generated so far,
it prefers solutions for new litmus tests to agree with previously generated rules.
This is \textit{not} only because avoiding conflicts saves time.
Intuitively, as \depsynth generates rules for increasingly more tests in the input set,
its confidence in the rules it has generated thus far becomes greater.
This gives further reason for \depsynth to prefer graphs that obey previously generated rules.

To give intuition about our second heuristic, newer writes tend to depend on older writes
because dependencies are "happens before" relationships.
From our experience, data updates are more likely to rely on previous updates than future ones.
As a result, prioritizing our search in this way leads to generating correct rules more often.
Of course, correct dependency graphs may not follow either of these heuristicts,
but since \sccsearch can backtrack on ordering decisions,
these heuristics do not compromise correctness.

\paragraph{Optimizations in \multisearch}
\autoref{alg:multi-search} describes a search over graphs for multiple litmus
tests that, like \autoref{alg:remove-edges}, removes edges one-at-a-time
to arrive at candidates. In practice, \depsynth also the uses SCC-splitting
optimization from \autoref{alg:scc-search} in \multisearch.

% Talk about using input rules as necessary paths?
% Additionally, \depsynth leverages necessary paths to prune the
% search away from graph sets that conflict with input rules.
% Rules output by \multisearch should not conflict with
% the set of input rules. Like necessary paths,

\paragraph{Skipping Tests}
Since \autoref{alg:top-level} gradually builds up a set of rules,
\depsynth will likely encounter litmus tests along the search where
the current candidate set of rules is sufficient to make the litmus
test crash consistent. When this is the case, \depsynth skips \rulessearch
for that test and does not generate any new rules. \depsynth still updates
the mapping from litmus tests to rules with the set of rules used for the test.
This is because the test may be relevant for later conflict resolution.

\paragraph{Reordering Tests}
Depending on the order of the input litmus tests, \depsynth may encounter tests for which
\rulessearch performs slowly. Specifically, when \rulessearch times out
for such a test, \depsynth postpones the search for the test until the end of the loop.
By doing so, \depsynth may be able to synthesize sufficient rules for the test by
finding rules for other, simpler tests. If so, \depsynth can skip the search for the
challenging test at the end.

Executing \rulessearch to a timeout is still expensive, and \depsynth avoids doing so
when possible by ordering the input litmus tests before performing \toplevel.
Specifically, \depsynth orders litmus tests based on the number of disk writes the
test generates, from least to greatest. This allows \depsynth to find rules for simpler
tests before moving on to more complex tests.

\fi