### 2023-04-17 chapters/depsynth/intro.tex

\section{Introduction}

% crash consistency is hard. non-determinism, reordering, consistency.
% it's all about ordering. a high-level consistency property but low-level writes.

% hard for two reasons:
% 1. requires reasoning about non-determinism and volatile state. so either very careful global reasoning
%    about what might be inflight at any point in the program (soft updates) or a centralized consistency
%    mechanism like a journal that sacrifices performance to make consistency easier.
% 2. the storage API provides no support for implementing consistency -- only provides durability
%    mechanisms like flush, that have to be used indirectly to get consistency guarantees.
% while some prior work improves this, still fundamentally requires the programmer to determine valid
% and invalid orderings at the granularity of individual writes.

% this paper: a new programming model for building storage systems in which a program synthesizer automatically
% infers the necessary ordering(?) to guarantee crash consistency.
% two parts:
% 1. instead of programming directly against a low-level disk API, programmers instead build against
%    a higher-level *ordering-aware buffer cache* API. requires programmer to label their writes to disk
%    and then configure the cache with a set of *dependency rules* to realize consistency
% 2. writing those rules is hard, so we develop a synthesizer than can automatically infer the rules
%    necessary to guarantee a desired consistency property. synthesized from a set of examples/litmus
%    tests. great because we can adapt to changes in functionality and in hardware.

% evaluated on a production storage system. we show that we can automatically generate dependency
% rules equivalent to the hand-written imperative implementation. also show that we can adapt to
% changes by changing the hardware. maybe we also implement a simple journaling FS.

Many applications build on storage systems
such as file systems and key-value stores
to reliably persist user data
even in the face of full-system crashes (e.g., power failures).
Guaranteeing this reliability requires the storage system to be \emph{crash consistent}:
after a crash, the system should recover to a consistent state without losing previously persisted data.
The state of a storage system is consistent if it satisfies the representation 
invariants of the underlying persistent data structures (e.g., a free data block must not be linked by any file's inode).
Crash consistency is notoriously difficult to get right~\cite{yang:explode,pillai:appcrash,zheng:torturing-db}, 
due to performance optimizations in modern software and hardware
that can reorder writes to disk
or hold pending disk writes in a volatile cache.
In normal operation, these optimizations are invisible to the user,
but a crash can expose their partial effects, leading to 
inconsistent states.

A number of general-purpose approaches exist to implement crash consistency,
including journaling~\cite{prabhakaran:journaling},
copy-on-write data structures~\cite{rodeh:btrfs},
and soft updates~\cite{ganger:soft-updates}.
However, implementing a storage system using these approaches is still challenging for two reasons.
First, practical storage systems combine crash consistency techniques
with optimizations such as log-bypass writes and transaction batching to improve performance~\cite{tweedie:ext2journal}.
These optimizations and their interactions are subtle,
and have led to severe crash-consistency bugs
in well-tested storage systems~\cite{lu:fsstudy,chen:dfscq}.
Second, developers must implement their system 
using low-level APIs provided by storage hardware and kernel I/O stacks,
which offer no direct support for enforcing consistency properties.
Instead they provide only durability primitives such as flushes,
and require the developer to roll their own consistency mechanisms on top of them.
While prior work offers testing~\cite{mohan:crashmonkey,yang:explode}
and verification~\cite{chen:fscq,sigurbjarnarson:yggdrasil} tools for validating crash consistency,
these tools do not alleviate the burden of implementing crash-consistent systems.

This paper presents a new synthesis-aided programming model for building crash-consistent storage systems.
% that \emph{automatically infers} the necessary write-level orderings to guarantee a desired crash consistency property.
The programming model consists of three parts: 
%
a high-level storage interface based on \emph{labeled writes};
%
a synthesis engine for turning labeled writes and a desired crash consistency property 
into a set of \emph{dependency rules} that writes to disk must respect; and 
%
a \emph{dependency-aware buffer cache} that enforces the synthesized rules at run time.
%
Together, these three components let developers keep their implementation free of 
hardcoded optimizations and mechanisms for enforcing consistency. 
%
Instead, developers can focus on the key aspects of their storage system---functional 
correctness, crash consistency, and performance---one at a time. 
%
Their development workflow consists of three steps.

First, developers implement their system against a higher-level storage interface
by providing \emph{labels} for each write their system makes to disk.
Labels provide information about the data structure the write targets
and the context for the write (e.g., the transaction it is part of).
For example, a simple journaling file system might require two writes to append to the journal:
one to append the data block to the tail of the journal (labeled \textsf{data})
and one to update a superblock that records a pointer to that tail (labeled \textsf{superblock}).
This higher-level interface allows the developer to assume a stronger
\emph{angelic nondeterminism} model for crashes---%
the system promises that crash states will \emph{always} satisfy
the developer's crash consistency property if possible---%
simplifying the implementation effort.\tighten

Second, to make their implementation crash consistent even on relaxed storage stacks,
the developer uses a new program synthesizer, \depsynth,
to automatically generate \emph{dependency rules} that writes to disk must respect.
A dependency rule uses labels to define an ordering requirement between two writes:
writes with one label must be persisted on disk before corresponding writes with the second label.
The \depsynth synthesizer takes three inputs: the storage system implementation,
a desired \emph{crash consistency predicate} for disk states of the storage system
(i.e., a representation invariant for on-disk data structures),
and a collection of small \emph{litmus test} programs~\cite{alglave:litmus-tool,bornholt:ferrite}
that exercise the storage system.
Given these inputs, \depsynth searches a space of happens-before graphs
to automatically generate a set of dependency rules
that guarantee the crash-consistency predicate for every litmus test.
% The developer can then supply these generated rules to their storage system.
Although this approach is example-guided and so only guarantees crash consistency on the supplied tests,
the dependency rule language is constrained to make it difficult to overfit to the tests,
and so in practice the rules generalize to arbitrary executions of the storage system.

Third, developers run 
their storage system on top of a \emph{dependency-aware buffer cache}
that enforces the synthesized dependency rules.
For example, in a journaling file system,
the superblock pointer to the tail of the journal must never refer to uninitialized data.
\depsynth will synthesize a dependency rule enforcing this consistency predicate
by saying that data writes must happen before superblock writes.
At run time, the dependency-aware buffer cache
enforces this rule by delaying sending writes labeled \textsf{superblock}
to disk until the corresponding \textsf{data} write has persisted.
The dependency-aware buffer cache is free to reorder writes in any way
to achieve good performance on the underlying hardware
(e.g., by scheduling around disk head movement or SSD garbage collection)
as long as it respects the dependency rules.

We evaluate the effectiveness and utility of \depsynth
in a case study that applies it to a production key-value store
developed by a major commercial cloud provider.%
\footnote{Anonymized for double-blind review.}
We show that \depsynth can rapidly synthesize dependency rules for this storage system.
By comparing those rules to the key-value store's existing crash-consistency behavior,
we find that \depsynth achieves similar results to rules hand-written by experts,
and even corrects an existing crash-consistency issue in the system automatically.
We also show that dependency rules synthesized by \depsynth
generalize beyond the example litmus tests used for synthesis,
and that \depsynth can be used for storage systems beyond key-value stores.
% Finally, we show that an automated crash consistency approach
% supports rapid iteration by using \depsynth
% to \emph{re}-synthesize dependency rules
% after changing the underlying storage hardware (and thus the guarantees it offers).

In summary, this paper makes three contributions:
\begin{itemize}
\item A new programming model for building storage systems that automates the implementation of crash consistency guarantees;
\item \depsynth, a synthesis tool that can infer the dependency rules sufficient for a storage system to be crash consistent; and
\item An evaluation showing that \depsynth supports different storage system designs and scales to production-quality systems.
\end{itemize}

\noindent
The remainder of this paper is organized as follows.
\cref{s:overview} gives a walk-through of building a simple storage system with \depsynth.
\cref{sec:problem} defines the \depsynth programming model, including labeled writes and dependency rules.
\cref{sec:alg} describes the \depsynth synthesis algorithm for inferring dependency rules,
and \cref{s:impl} details \depsynth's implementation in Rosette.
\cref{sec:eval} evaluates the effectiveness of \depsynth.
\cref{s:related} discusses related work, and \cref{s:conclusion} concludes.\tighten

\if 0
Modern storage systems rely on crash consistency to maintain data across disk crashes.
As the number of disks used by these systems surpasses the order of hundreds and thousands,
the likelihood of any single disk crashing increases tremendously, thus making crash
consistency even more important. However, reasoning about the crash consistency of systems
is not an easy task. Moreover, not all methods of achieving crash consistency are created equal.
Journaling introduces slowdowns in the system due to duplicating writes.
Systems using soft updates avoid write duplication by instead
ordering writes sent to disk in such a way that all intermediate disk states are consistent.
The downside of soft updates lies in the difficulty of producing such a correct order.

Soft updates have been used in a variety of storage systems in
order to ensure crash consistency \todo{cite}. To implement soft updates, system
developers must correctly order disk writes in such a way that if a disk crash occurs
at any point, the disk will be left in a consistent state. As shown in \todo{FeatherStitch},
one way to keep track of the correct order of disk writes is to encode a set of \textit{dependencies}
between individual disk writes. The disk is allowed to order writes in any way that
obeys this set of dependencies.
Since some orders can be executed more quickly than others, it is desirable to minimize
write dependencies to only those necessary to ensure crash consistency.

This paper presents a tool for automatically implementing crash consistency through soft updates
without requiring the developer to reason about the crash states of their system. We develop
\depsynth, a framework for synthesizing \textit{dependency rules} that generate write dependencies
for any execution of the input system. Using \depsynth, we implement a key-value store based
on a simplified version of \todo{ShardStore}. \depsynth synthesizes dependency rules for this
system that generate dependencies similar to those hand-written by experts.
% It would be nicer if I could say that we synthesized rules for several simpler key-value stores as well

The dependency rules output by \depsynth are constrained by three main requirements. First, the output
rules must ensure crash consistency for any execution of the system. More specifically, any
order of disk writes allowed by the rules output from \depsynth should never put the system in an
inconsistent state. Second, \depsynth should output as few rules as possible in order to satisfy the first
constraint. The reason for this requirement is that some executing schedules for disk writes are faster than
others. Ideally, the system has the freedom to choose any valid schedule of disk writes to execute. Third,
the rules output by \depsynth should be acyclic. This means that the rules will never generate dependencies
such that two disk writes depend on each other. This is important because the system may not be able to enforce
that two writes are fully written to the disk at the same time. The next few paragraphs explain how \depsynth
meets all three of these requirements.
\todo{Second-to-last sentence feels weird. How to better explain the third requirement?}

% TODO explain how \depsynth meets these three requirements
The first goal of \depsynth is to output rules that are sufficient for any execution over the system. One way to
achieve this goal would be to enumerate candidate sets of rules and check with an SMT solver that
the rules are sufficient for any reachable disk state. However, for realistic storage systems, the space of
reachable states is too large and complex for solvers to reason about efficiently. Instead of considering
all reachable states of a system, we take a finite set of \textit{concrete} executions as part of the
specification. \depsynth then synthesizes a set of dependency rules sufficient for all states reachable
in the set of specified executions. Since our goal is for generated rules to be sufficient for all
possible executions, we have designed the language for dependency rules in such a way that rules
generated by \depsynth generalize to executions outside of the input set.

The second goal is that output rules should be \textit{minimal} in the number of generated dependencies
so that the storage system has the freedom to execute the most optimal valid schedule. \depsynth achieves this
goal by splitting the rule search for a single litmus tests into two phases: graph generation and rule
extrapolation. In the first phase, \depsynth generates a minimal dependency graph for the litmus test
using an enumerative backtracking search. The second phase extrapolates the minimal dependency graph
into a set of rules that generates the graph.
% Importantly, the set of rules may generate dependencies that are not in the graph...

The final goal is to generate a set of rules that is \textit{acyclic} so that any output dependency graph
can be enforced during an actual execution as a partial order over the writes in the storage system.
This goal is achieved through both a local backtracking strategy in the per-litmus-test graph search
as well as the global search for rules over all litmus tests. Backtracking in the local search is
relatively straightforward: the enumerative search over graphs for a single litmus test should backtrack
to avoid cycles in the graph. For the global search, since rules for each litmus test are combined for a
final output, \depsynth may
generate conflicting rules over multiple litmus tests that only cause a cycle when combined.
When this happens, \depsynth resolves the conflicts by generating new rules for all
involved litmus tests.

To summarize, this paper makes the following contributions:
\begin{enumerate}
  \item \depsynth, the first tool for synthesizing crash consistency code for storage systems
        in the form of soft updates.
  \item A new formalization of the \textit{dependency rule synthesis problem}. % TODO elaborate
  \item A novel backtracking search that enables \depsynth to solve the dependency rule synthesis problem.
  \item An evaluation that shows \depsynth can efficiently generate a set of rules that generalizes
        to outputs outside the training set and that is comparable to expert-written crash consistency code.
\end{enumerate}

The rest of this paper is organized as follows.
\autoref{s:overview} illustrates how developers use \depsynth to construct a basic crash consistent storage system.
\autoref{s:problem} formalizes the dependency rule synthesis problem.
\autoref{s:algorithm} presents the \depsynth algorithm for searching for rules over a set of litmus tests.
\autoref{s:impl} provides implementation details.
\autoref{s:eval} evaluates \depsynth.
\autoref{s:related} discusses related work.
\autoref{s:conclusion} concludes.
\fi