### 2023-04-17 figs/alg.tex

\begin{algorithmic}[1]
\Function{\depsynthalg}{\impl, \tests, \consist}
\State $\ruleset \gets \{\}$
\Loop
    \State $\test \gets \Call{NextTest}{\tests, \ruleset, \impl, \consist}$  \label{alg:depsynth:nexttest}
    \If{$\test = \bot$}  \Comment{\ruleset makes all tests in \tests crash consistent}
        \State \Return \ruleset
    \EndIf
    \State $\tests \gets \tests \setminus \test$
    \State $\ruleset' \gets \Call{\rulesfortest}{\test, \impl, \consist}$ \label{alg:depsynth:rulesone}
    \If{$\ruleset' = \bot$}
        \State \Return \UNSAT  \Comment{No rules can make \test crash consistent}
    \EndIf
    \State $\ruleset \gets \ruleset \cup \ruleset'$ \label{alg:depsynth:add}
    \If{$\neg \Call{Acyclic}{\ruleset}$}  \Comment{Fail if new rules create a cycle in the rule set}
        \State \Return \UNKNOWN  \label{alg:depsynth:resolve}
    \EndIf
\EndLoop
\EndFunction

\vspace{0.8em}

\Function{NextTest}{\tests, \ruleset, \impl, \consist}
    \For{$\test \in \tests$}
        \If{$\neg \Call{\crashconsistentalg}{\test, \ruleset, \impl, \consist}$}
            \State \Return \test
        \EndIf
    \EndFor
    \State \Return $\bot$
\EndFunction

\vspace{0.8em}

\Function{\crashconsistentalg}{$\test = \langle P_\emph{initial}, P_\emph{main} \rangle$, \ruleset, \impl, \consist}  \Comment{Check Def.~\ref{def:crash-consistent} with an SMT solver}  \label{alg:depsynth:cc}
    \State $d_\emph{initial} \gets \run{\evaluate{P_\emph{initial}}{\mathcal{O}}}{\vec{\top}}{d_0}$
    \State \Return $\forall \vec{s}.\ \valid{\vec{s}}{P_\emph{main}}{R} \Rightarrow \consistent{\run{\evaluate{P_\emph{main}}{\mathcal{O}}}{\vec{s}}{d_\emph{initial}}}$
\EndFunction

\algstore{depsynth}
\end{algorithmic}