### 2023-04-17 figs/resolve.tex

\newcommand{\impl}{\ensuremath{\mathcal{O}}\xspace}
\newcommand{\test}{\ensuremath{T}\xspace}
\newcommand{\tests}{\ensuremath{\mathcal{T}}\xspace}
\newcommand{\consist}{\ensuremath{\emph{Consistent}}\xspace}
\newcommand{\rulemap}{\ensuremath{\mathcal{R}}\xspace}
\newcommand{\ruleset}{\ensuremath{R}\xspace}
% \newcommand{\writesused}{\ensuremath{W_U}\xspace}
% \newcommand{\writesleft}{\ensuremath{W_R}\xspace}
\renewcommand{\wr}{\ensuremath{W}\xspace}
\newcommand{\ord}{\ensuremath{V}\xspace}
\newcommand{\gr}{\ensuremath{G}\xspace}
\newcommand{\comp}{\ensuremath{C}\xspace}

\begin{algorithmic}[1]
\algrestore{depsynth}

\Function{ResolveCycle}{\rulemap, \impl, \consist}
    \State $C \gets \Call{GetCycle}{\cup_{(T' \mapsto R) \in \rulemap} R}$  \Comment{Identify rules $C \subseteq R$ that participate in the cycle}
    \State $\tests' \gets \{ T \ | \ (T \mapsto R) \in \rulemap \land T \in C \}$  \Comment{Collect the tests that generated rules in $C$}
    \State \Return \Call{RulesForTests}{\tests', \impl, \consist}
\EndFunction

\end{algorithmic}