### 2023-04-17 thesis.tex

\documentclass[12pt]{book}

%% Added by Jacob
\usepackage{makecell}
%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}

\usepackage{algorithm}
\usepackage{algpseudocode}
\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\algnewcommand\Not{\textbf{not} }
\newcommand{\algorithmautorefname}{Algorithm} % autoref

\usepackage{adjustbox}
\usepackage[headings]{fullpage}
\usepackage{xargs}
\usepackage{xspace}
\usepackage[doublespacing]{setspace}

\usepackage{microtype}
\usepackage[T1]{fontenc}

\usepackage{wrapfig}
\usepackage[font={sf}]{caption}
\usepackage[subrefformat=parens]{subcaption}

\usepackage{multirow}

\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{{_build/}{figures/}}

\usepackage{amsmath}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{proof}{Proof}[chapter]

\setlength{\marginparwidth}{2cm}
\usepackage[textsize=tiny]{todonotes}
%\renewcommand\max[2][]{\todo[color=orange, #1]{\sffamily #2}}

\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage[scaled=.85]{DejaVuSansMono}

%\usepackage{listings}
\usepackage[procnames]{listings}          % format code
\lstset{
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  stringstyle=\slshape\color{green!40!black},
  basicstyle=\ttfamily\small,
  language=Python,
  morekeywords={*, self},
  commentstyle=\slshape\color{black!60},
  tabsize=2,
}

\lstdefinelanguage{Rust}{
  sensitive,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  moredelim=[s][{\itshape\color[rgb]{0,0,0.75}}]{\#[}{]},
  morestring=[b]{"},
  alsodigit={},
  alsoother={},
  alsoletter={!},
  % keywords
  otherkeywords={=>},
  morekeywords={break, continue, else, for, if, in, loop, match, return, while},
  morekeywords={as, const, let, move, mut, ref, static, unsafe},
  morekeywords={dyn, enum, fn, impl, Self, self, struct, trait, type, use, where},
  morekeywords={crate, extern, mod, pub, super},
  morekeywords={abstract, alignof, become, box, do, final, macro,
    offsetof, override, priv, proc, pure, sizeof, typeof, unsized, virtual, yield},
  % traits
  morekeywords=[2]{Send},
  % types
  morekeywords=[3]{bool, char, f32, f64, i8, i16, i32, i64, isize, str, u8, u16, u32, u64, unit, usize, i128, u128},
}%

\def\mytitle{Program Synthesis for Systems Developers}
\def\myauthor{Jacob Van Geffen}
\def\year{2023}

\title{\mytitle}
\author{\myauthor}

% TODO write thesis statement here
\newcommand\thesisstmt{%
  e-graphs and equality saturation are compelling techniques for
  program representation and manipulation that should now be considered
  for programming tools across many domains.
}

\newcommand\Thesisstmt{\expandafter\MakeUppercase \thesisstmt}

\input{macros}
\input{dmacros}
\input{data/nr}

\begin{document}
\pagestyle{empty}

% title and copyright pages
\begin{center}
  {\huge \mytitle}
  \vfill

  {\Large \myauthor}
  \vfill

  \begin{spacing}{1}
    A dissertation \\
    submitted in partial fulfillment of the \\
    requirements for the degree of
  \end{spacing}
  \vfill

  Doctor of Philosophy
  \vfill

  University of Washington \\
  \year
  \vfill

  Reading Commitee: \\
  Emina Torlak, Chair \\
  Xi Wang \\
  Zachary Tatlock \\
  Shane Steinert-Threlkeld
  \vfill

  \begin{spacing}{1}
    Program Authorized to Offer Degree: \\
    \cse
  \end{spacing}
  \clearpage

  \textcopyright{} Copyright \year\\
  \myauthor
  \clearpage
\end{center}

\pagestyle{plain}
\setcounter{page}{1}
\pagenumbering{roman}

% abstract page
\begin{center}
  University of Washington \\[1em]
  \textbf{Abstract}        \\[1em]
  \mytitle                 \\[1em]
  \myauthor                \\[1em]

  % Supervisory Committee: \\[-0.5em]
  % Luis Ceze, Chair \qquad
  % Adriana Schulz   \qquad
  % Zachary Tatlock  \\[-0.5em]
  Chair of the Supervisory Committee: \\[-0.5em]
  Emina Torlak \\[-0.5em]
  \pgas
  \\[2em]
\end{center}
\input{abstract}

\begin{spacing}{1.5}
  \tableofcontents
\end{spacing}

% TODO write acknowledgements HERE
\chapter*{Acknowledgments}

This thesis covers work done in the final year and a half of my PhD.
The work done before then,
 largely with the Molecular Information Systems Lab at UW,
 was and is a great source of pride.
I am very fortunate to enjoy the
 endlessly supportive mentorship of my advisor Luis Ceze,
 who is the very first person you want on your team,
 and who introduced me to
 the collaborative, cross-disciplinary environment at MISL.

On the \egg side of things,
 I owe the all the fun and any success
 I had to the entire PLSE group.
Thanks especially to Zach Tatlock,
 without whom PLSE would simply not be the same.
The main component of this thesis is the \egg
 paper, and my co-authors
 in that work deserve a special shoutout for making
 that project what it is:
 Chandrakana Nandi,
 Remy Wang,
 Oliver Flatt,
 Zach Tatlock,
 and Pavel Panchekha.

Thanks to my friends and family for
 making this journey possible and even enjoyable.
And finally,
 thank you Sami for doing this whole thing with me.

\clearpage

\pagestyle{headings}
\setcounter{page}{1}
\pagenumbering{arabic}
\renewcommand{\chaptermark}[1]{\markboth{\sc{\chaptername\ \thechapter.\ #1}}{}}
\renewcommand{\sectionmark}[1]{\markright{\sc{\thesection.\ #1}}{}}
% \listoftodos
%\input{chapters/01-intro}
%\input{chapters/02-background}
%\input{chapters/03-rebuilding}
%\input{chapters/04-analyses}
%\input{chapters/05-egg}
%\input{chapters/06-case-studies}
%\input{chapters/99-conclusion}
\input{chapters/depsynth/depsynth}

\clearpage
\addcontentsline{toc}{chapter}{Bibliography}
\singlespacing
\bibliographystyle{alpha}
\bibliography{references}

\end{document}