### 2023-05-04 chapters/intro.tex
Intro first draft

\chapter{Introduction}

% What is hard?
In recent years, the development of software systems has become increasingly
complex.
From the construction of large production key-value stores to Linux kernel JIT compilers
% TODO should I include these examples?
to the design of blockchain protocols and fault-tolerent networks,
modern systems must satisfy ever-increasing requirements on performance.
Additonally, the
correctness of these systems is now more important than ever. The
afformentioned storage systems are responsible for safely storing the data
of millions of people, while
the incorrect behavior of in-kernel compilers can subject the tens of millions of
Linux users to a slew of vulnerabilities.

% Why is it hard?
Modern systems developers must tackle three challenges when constructing
systems that both meet high performance criteria and exhibit correct behavior.
First, developers must \textbf{specify} the intended behavior of their system at a high
level. This specification can be encoded in a variety of ways, from
English-text design documents to formal logic specifications.
Second, developers must \textbf{implement} their system in accordance with the defined
specificaiton. Generally, this implementation done at a lower level than the
specification and may contain details or optimizations not referenced by the specification.
Third, developers must \textbf{verify} that their implementation behaves correctly 
according to their specification. Much like the last two tasks, verification can be done
in one of several ways, from testing to manual or automated proofs to code review.
However, no matter the method for each of these three tasks,
all pose a significant burden of effort to systems developers.

% In order to ameliorate the burden \todo{continue} Various areas of work have
% aimed to limit the burdens ,

% How can we make it easier?
Between these three efforts, system developers repeat encoding the semantics of
their system: the implementation of the system must adhere to the semantics
given by the specification, and manual verification efforts like testing again
repeat these semantics for specific inputs to the system.
The goal of my work is to take advantage of the overlap between these three tasks
and reduce the burden on systems developers with \textit{program synthesis}.
Program synthesis is an automated programming technique
that takes as input a specification and outpus both an implementation
as well as some certificate verifying that the implementation correctly behaves
according to the input specification.
This work presents new techniques that enable the construction of program
synthesis tools for systems software, and additionally demonstrates the
efficacy of these tools.

%In the context of 
%By doing so, synthesis eliminates the
%effort from developers to both implement and verify their systems, requiring
%only a specification for the system.

% What will I say to demonstrate the "how"? 
%In this work, I aim to
%demonstrate the benefits of using program synthesis tools in order to both
%design and implement systems software.

%This work presents new techniques that enable the construction of program
%synthesis tools for systems software, and additionally demonstrates the
%efficacy of these tools.
My thesis is that \textit{program
synthesis is an effective tool for designing systems that are both correct and
performant}.  This dissertation explores this thesis over two particular types
of systems --- just-in-time compilers and crash consistent storage systems ---
and demonstrates how new synthesis tools can reduce implementation and verification
effort for developers despite the size and complexity of these systems.

\section{Challenges}

The main challenge in building any synthesis tool comes from the size of the
search space of possible implementations.
In order to generate verifiably correct implementations, synthesis tools both limit
their search space and order their search to prioritize the most likely candidates.
One common technique synthesis tools use to both limit and order the search space is
to specialize over a particular domain.
For example, a large body of work focuses solely on synthesizing classes of spreadsheet
table transformations \todo{cite}.
These tools restrict their space to sequential programs that invoke a set
of operations over relevant to these particular transformations, and also intelligently
order their search to prioritize candidates that are most likely desired by the synthesis
tool user.
% By doing so, % TODO finish this?
Another common technique for ordering the space of candidate output programs is to
consider the size of candidates.
Since the number of candidate programs increases exponentially with size,
synthesis tools often prioritize searching for smaller programs before larger ones
(EXAMPLES\todo{Cite examples} for example).
In general, synthesizing large complex programs is much more difficult than
synthesizing small blocks of code.

Unline most existing target domains for program synthesis, systems software reasons
about multiple layers of abstraction.
Specifically, many systems consider domain-specific language (DSL) programs as input,
so synthesis tools for systems must reason about the execution of both input DSL programs
and the system implementation itself.
%these large systems often encode highly complex semantics
%which often reason at multiple levels of abstraction.
Additionally, systems software
often requires a monolithic amount of code, on the order of tens of thousands
of lines \todo{cite examples of big systems}.
%manually generated proofs of correctness for these systems can be just as large if not
%moreso. \todo{cite big proofs of verified OS?}
This means that na\"ively applying existing techniques for program synthesis to generate
systems software is infeasible.

In this work, I identify two key observations that enable the creation of synthesis tools
that effectively generate correct and performant systems software.
First, \textbf{program decomposition} allows synthesis tools to generate smaller target programs.
By considering systems as a conjunction of smaller pieces,
synthesis tools can search for candidate programs that implement a single piece.
The resulting synthesized program can be combined with either user-written code or
other synthesized system pieces, effectively reducing the developer burden.
% JitSynth: break up into minicompilers (TODO)
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
Second, system-specific \textbf{metaprogram abstractions} enable synthesis tools to reason 
about candidate programs that themselves take DSL programs as input.
By reasoning about these simpler abstractions rather than the complex systems directly,
synthesis tools are able to more efficiently search over candidate programs.
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)
In the next two sections, I describe how these observations enabled the creation of
synthesis tools for both in-kernel JIT compliers and for crash consistent storage systems.

%%%%%%%%%% %%%%%%%%%% %%%%%%%%%% %%%%%%%%%% %%%%%%%%%% %%%%%%%%%% %%%%%%%%%% %%%%%%%%%%

\section{Synthesis for In-Kernel Compilers}
% TODO
% JitSynth: break up into minicompilers (TODO)
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)

\section{Automatic Crash Consistency with Synthesis}
% TODO
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)

\section{Contributions}