### 2023-05-05 chapters/jitsynth/rebuttal.txt
acks, jitsynth, dmacros

We thank the reviewers for their insightful feedback. We will use it to revise
the paper. Below, we answer the questions from each of the reviews.

---R1---

> Future research in speeding up synthesis could very likely complement the
  approaches presented here. ... Perhaps combining these techniques with
  peephole superoptimizers would help.

Yes, we also believe that these are promising directions for future work, and
that they can help improve the performance of both the synthesis process and the
synthesized JITs.

> It would be great if [the code] were [publicly] available.

We plan to submit an artifact if the paper is accepted. We will
also make the code and all of our benchmarks publicly available.

---R2---

> Why [87 out of 102 eBPF instructions] enough for the compiler to
  be useful? ... How many real eBPF programs require just those 87 instructions
  (how many of your test programs used an instruction you didn't handle?)

These 87 instructions cover 228/244 (~93%) of the eBPF programs in the
test suite in the Linux kernel (lib/test_bpf.c). They are also sufficient
to run all the application benchmarks from the Jitk project (Fig. 7). We
will clarify this in the revision.

> Page 12 [says] that the sketch for addi32 includes 2^350 programs,
  but ... that number seems far too low.

Yes, 2^350 is much lower than the number of all programs at this tree depth and
program length. We got this lower number by including some simple optimizations
in the naive sketch, such as reusing symbolic constants.

> Bugs in Definitions 8 and 9.

Yes, these are indeed bugs in the definitions; thanks for catching them!  The
implementation matches the intuition provided in the text, and we will fix the
definitions accordingly.

> Is strong soundness equivalent to soundness?  Are there any compilers that are
  sound but not strongly sound? It looks ... like every sound compiler is
  strongly sound.

This is a great question. While the logical formulas for soundness and strong
soundness aren't equivalent, we could not think of any case in which a compiler
is sound but not strongly sound. For that reason, we also strongly suspect that
using strong soundness won't miss (m)any real compilers.

> A 2.28x slowdown is fairly significant, so it's a bit strong to say
that "the performance is similar to the existing Linux compiler".

Agreed, we will tone down this claim. This kind of slowdown would be okay
for some eBPF programs (e.g., for system-call filtering), but not for
high-performance network filtering. For the latter, it would be necessary
to get the performance on par with that of the hand-written compiler.

> [Did you try] to synthesise compilers that don't use bitvector constants
  ... or that use small constants ... as one of your intermediate sketches?

> [Did you try] to generate mini-compilers for common sequences of
  instructions, or ... a peephole optimiser for some common constructs?

These are great suggestions. While we haven't tried them, we believe that they
would improve synthesis performance as well as the performance of the generated
code.

---R3---

Thanks for the review! We agree that more work is needed before automatically
generated JITs can replace the ones written by hand.

---R4---

> How easy is it to extend the compiler to handle dynamic register allocation?

This would be quite difficult, as it breaks our core assumption that a static
register mapping exists from the source to target language, which is a common
practice for in-kernel JITs.

> You mentioned a way to use pre-defined micros/subroutines as a way to simplify
  compilation of commonly used and/or complex instructions. Should these
  micros/subroutines be verified for correctness and security?

These pre-defined macros are also synthesized by JitSynth, and they are verified as
part of the synthesis process. We will clarify this in any future revisions.

>  It would be interesting to compare and combine this work with Ironclad Apps
   ... and [allow] the compiler developer to interact with the synthesis tool and
   guide the generation.

We agree that these are interesting directions for future work; thanks for the
suggestions!