### 2023-05-05 chapters/jitsynth/jitsynth.tex

\chapter{JitSynth: Synthesizing Just-In-Time Compilers for In-Kernel DSLs}
\label{c:jitsynth}
% \input{macros}

%\begin{abstract}

Modern operating systems allow user-space applications to submit code for kernel
execution through the use of in-kernel domain specific languages (DSLs).
Applications use these DSLs to customize system policies and add new
functionality. For performance, the kernel executes them via just-in-time (JIT)
compilation. The correctness of these JITs is crucial for the security of the
kernel: bugs in in-kernel JITs have led to numerous critical issues and
patches.\tighten

\smallskip
This paper presents \jitsynth, the first tool for synthesizing verified JITs for
in-kernel DSLs. \jitsynth takes as input interpreters for the source DSL and the
target instruction set architecture. Given these interpreters, and a mapping
from source to target states, \jitsynth synthesizes a verified JIT compiler from
the source to the target. Our key idea is to formulate this synthesis problem as
one of synthesizing a per-instruction compiler for \emph{abstract register
machines}. Our core technical contribution is a new \emph{compiler metasketch}
that enables \jitsynth to efficiently explore the resulting synthesis search
space. To evaluate \jitsynth, we use it to synthesize a JIT from eBPF to RISC-V
and compare to a recently developed Linux JIT\@. The synthesized JIT avoids all
known bugs in the Linux JIT, with an average slowdown of
$\EbpfCompilerSlowdown\times$ in the performance of the generated code. We also
use \jitsynth to synthesize JITs for two additional source-target
pairs.  The results show that \jitsynth offers a promising new way
to develop verified JITs for in-kernel DSLs.\tighten

%\end{abstract}

\section{Introduction}
\input{chapters/jitsynth/intro}

\input{chapters/jitsynth/overview}

\input{chapters/jitsynth/problem}

\input{chapters/jitsynth/search}

\input{chapters/jitsynth/impl}

\input{chapters/jitsynth/results}

\input{chapters/jitsynth/related}

\input{chapters/jitsynth/conclusion}

%\bibliographystyle{splncs04}
%\bibliography{n-str,jitsynth,n,n-conf}

% \input{appendix}