### 2023-05-05 chapters/jitsynth/related.tex

\section{Related Work}\label{s:related}

\paragraph{JIT compilers for in-kernel languages.}

JIT compilers have been widely used to improve
the extensibility and performance of systems software,
such as OS kernels~\cite{chen:vmsec,engler:vcode,fleming:ebpf,myreen:vjit}.
%
One notable system is Jitk~\cite{wang:jitk}.
It builds on the CompCert compiler~\cite{leroy:compcert}
to compile classic BPF programs to machine instructions.
%
Both Jitk and CompCert are formally verified for correctness
using the Coq interactive theorem prover.  Jitk is further extended
to support eBPF~\cite{sobel:ejitk}.
%
Like Jitk,
\jitsynth provides formal correctness guarantees
of JIT compilers.
Unlike Jitk,
\jitsynth does not require developers to write either the implementation or proof
of a JIT compiler.
Instead,
it takes as input interpreters of both source and target languages
and state-mapping functions,
using automated verification and synthesis to produce a JIT compiler.

An in-kernel extension system such as eBPF
also contains a \emph{verifier},
which checks for safety and termination of input programs~\cite{gershuni:crab-ebpf,wang:jitk}.
\jitsynth assumes a well-formed input program that passes the verifier
and focuses on the correctness of JIT compilation.

\paragraph{Synthesis-aided compilers.}

There is a rich literature that explores generating and synthesizing
peephole optimizers and superoptimizers based on a given ISA or language
specification~\cite{bansal:superopt,davidson:peephole,gulwani:brahma,joshi:denali,massalin:superopt,sasnauskas:souper,schkufza:stoke}.
%
Bansal and Aiken described a PowerPC-to-x86 binary translator using
peephole superoptimization~\cite{bansal:binary}.
%
Chlorophyll~\cite{phothilimthana:chlorophyll} applied synthesis
to a number of compilation tasks for the GreenArrays GA144 architecture,
including code partitioning, layout, and generation.
%
\jitsynth bears the similarity of translation between a source-target pair
of languages and shares the challenge of scaling up synthesis.
%
Unlike existing work,
\jitsynth synthesizes a \emph{compiler} written in a host language,
and uses compiler metasketches for efficient synthesis.
%which is further translated into C for execution in an OS kernel.

\paragraph{Compiler testing.}
Compilers are complex pieces of software and are known to be difficult
to get right~\cite{marcozzi:compiler-fuzzing}.
Recent advances in compiler testing,
such as Csmith~\cite{yang:csmith} and EMI~\cite{zhang:emi},
have found hundreds of bugs in GCC and LLVM compilers.
%
Alive~\cite{lee:aliveinlean,lopes:alive} and Serval~\cite{nelson:serval}
use automated verification techniques to uncover bugs in the LLVM's peephole optimizer
and the Linux kernel's eBPF JIT compilers, respectively.
%
\jitsynth complements these tools by providing a correctness-by-construction
approach for writing JIT compilers.