### 2023-05-05 chapters/jitsynth/search.tex

\section{Solving the \MiniCompiler Synthesis Problem}
\label{s:algorithm}

This section presents our approach to solving the \minicompiler synthesis
problem defined in \autoref{s:problem}. We employ syntax-guided
synthesis~\cite{solar-lezama:sketch} to search for an implementation of a
\minicompiler in a space of candidate programs. Our core contribution is an
effective way to structure this space using a \emph{compiler metasketch}. This
section presents our algorithm for generating compiler metasketches, describes
its key subroutines and optimizations, and shows how to solve the resulting
sketches with an off-the-shelf synthesis engine.\tighten
%, and highlights key details of our implementation of \jitsynth.\tighten

\subsection{Generating Compiler Metasketches}

\jitsynth synthesizes \minicompilers by generating and solving
\emph{metasketches}~\cite{bornholt:synapse}. A metasketch describes a space of
candidate programs using an ordered set of syntactic templates or
\emph{sketches}~\cite{solar-lezama:sketch}. These sketches take the form of
programs with missing expressions or \emph{holes}, where each hole describes a
finite set of candidate completions. \jitsynth sketches are expressed in a
\emph{host language} \host that serves both as the implementation language for
\minicompilers and the specification language for ARMs. \jitsynth expects the
host to provide a synthesizer for completing sketches and a symbolic evaluator
for reducing ARM semantics to SMT constraints. \jitsynth uses these tools to
generate optimized metasketches for \minicompilers, which we call \emph{compiler
metasketches}.\tighten

\autoref{alg:compiler-metasketch-generator} shows our algorithm for generating
compiler metasketches. The algorithm, \CMS, takes as input an abstract source
instruction $\iota$ for a source machine $\arm_S$, a target machine $\arm_T$,
and a state mapping \Map from $\arm_S$ to $\arm_T$. Given these inputs, it
lazily enumerates an infinite set of \emph{compiler sketches} that collectively
represent the space of all straight-line bitvector programs from $\prog{\iota}$
to $\mathid{List}(\prog{\mathcal{I}_T})$. In particular, each compiler sketch
consists of $k$ target \emph{instruction holes}, constructed from {field holes}
that denote bitvector expressions (over the fields of $\iota$) of depth $d$ or
less. For each length $k$ and depth $d$, the \CMS loop generates three kinds of
compiler sketches: the \emph{pre-load}, the \emph{read-write}, and the
\emph{naive} sketch. The naive sketch (\autoref{s:algorithm:naive}) is the most
general, consisting of all candidate \minicompilers of length $k$ and depth $d$.
But it also scales poorly, so \CMS first yields the pre-load
(\autoref{s:algorithm:pld}) and  read-write (\autoref{s:algorithm:rw}) sketches.
As we will see later, these sketches describe a subset of the programs in the
naive sketch, and they are designed to prioritize exploring small parts of the
search space that are likely to contain a correct \minicompiler for $\iota$, if
one exists.

\begin{figure}[h]
\begin{algorithmic}[1] 
  \Function{CMS}
    {$\iota, \arm_S, \arm_T, \Map$} \Comment{$\iota\in\mathcal{I}_S, \arm_S = (\mathcal{I}_S, \ldots)$}
    \For{$n\in\mathbb{Z}^+$} \Comment{Lazily enumerates all compiler sketches}
      \For{$k \in [1, n], d = n-k$} \Comment{of length $k$ and depth $d$,}
        \State \textbf{yield} $\LCS(k,d,\iota,\arm_S,\arm_T,\Map)$ \Comment{yielding the pre-load sketch first,}
        \State \textbf{yield} $\RW(k,d,\iota,\arm_S,\arm_T,\Map)$ \Comment{read-write sketch next, and}
        \State \textbf{yield} $\Naive(k,d,\iota,\arm_S,\arm_T,\Map)$ \Comment{the most general sketch last.}
      \EndFor
    \EndFor
  \EndFunction
\end{algorithmic}
\caption{Compiler metasketch for the abstract source instruction
$\iota$, source machine $\arm_S$, target machine $\arm_T$, and state mapping
$\mathcal{M}$ from $\arm_S$ to
$\arm_T$.\tighten}\label{alg:compiler-metasketch-generator}
\end{figure}

\subsection{Generating Naive Sketches}\label{s:algorithm:naive}

The most general sketch we consider, $\Naive(k,d,\iota,\arm_S,\arm_T,\Map)$, is
shown in \autoref{fig:naive-sketch}. This sketch consists of $k$ instruction
holes that can be filled with any instruction from $\mathcal{I}_T$. An
instruction hole chooses between expressions of the form $(\mathit{op}_T, H)$,
where $\mathit{op}_T$ is a target opcode, and $H$ specifies the field holes for
that opcode. Each field hole is a bitvector expression (of depth $d$) over the
fields of the input source instruction and arbitrary bitvector constants. This
lets target instructions use the immediates and registers (modulo \Map) of the
source instruction, as well as arbitrary constant values and register names.
Letting field holes include constant register names allows the synthesized
\minicompilers to use target registers unmapped by \Map as temporary, or
scratch, storage. In essence, the naive sketch describes all straight-line
compiler programs that can make free use of standard C arithmetic and bitwise
operators, as well as scratch registers.\tighten

The space of such programs is intractably large, however, even for small inputs.
For instance, it includes at least $2^{350}$ programs of length $k=5$ and depth
$d\leq 3$ for the toy example from \autoref{s:overview}. \jitsynth therefore
employs two effective heuristics to direct the exploration of this space toward
the most promising candidates first, as defined by the read-write and pre-load
sketches.\tighten

\begin{figure}
\begin{algorithmic}[1] 
  \Function{Naive}
    {$k, d, \iota , \arm_S, \arm_T, \Map$} \Comment{$\iota\in\mathcal{I}_S, \arm_S = (\mathcal{I}_S, \ldots)$}
    \State $(\mathid{op}, \mathcal{F}) \gets \iota$, $(\mathid{I}_T, \ldots) \gets \arm_T$ \Comment{Source instruction, target instructions.}
    \State $p \gets \mathid{FreshId}()$ \Comment{Identifier for the compiler's input.}
    \State $\mathid{body}\gets []$  \Comment{The body of the compiler is a sequence}
    \For{$0 \leq i < k$} \Comment{of $k$ target instruction holes.}
      \State $I \gets \{\}$  \Comment{The set $I$ of choices for a target instruction hole}
      \For{$(\mathid{op}_T, \mathcal{F}_T)\in\mathcal{I}_T$} \Comment{includes all instructions from $\mathcal{I}_T$.} \label{li:instruction-hole}
        \State $E \gets \{\mathid{Expr}(p.f, \Map) \,|\, f\in\mathid{dom}(\mathcal{F})\}$ \Comment{Any source field can appear in}
        \State $H \gets \{ f\mapsto \mathid{Field}(\mathcal{F}_T(f),d,E) \,|\, f\in\mathid{dom}(\mathcal{F}_T)\} $ \Comment{a target field hole, and}\label{li:field-holes}
        \State $I\gets I \cup \{ \mathit{Expr}((\mathid{op}_T, H), \Map) \}$ \Comment{any constant register or value.}
      \EndFor
      \State $\mathid{body} \gets \mathid{body} \cdot [\mathid{Choose}(I)]$ \Comment{Append a hole over $I$ to the body.}
    \EndFor
    \State \Return $\mathid{Expr}((\lambda p\in\prog{\iota}\, .\, \mathid{body}),\Map)$ \Comment{A \minicompiler sketch for $\iota$.}
  \EndFunction
\end{algorithmic}
\caption{Naive sketch of length $k$ and maximum depth $d$ for $\iota$, $\arm_S$,
$\arm_T$, and \Map. Here, $\mathid{Expr}$ creates an expression in the host language, 
using \Map to map from source to target register names and memory addresses;
$\mathid{Choose}(E)$ is a hole that chooses an expression from the set $E$; and
$\mathid{Field}(\tau,d,E)$ is a hole for a bitvector expression of type
$\tau$ and maximum depth $d$, constructed from arbitrary bitvector constants and
expressions $E$.\tighten}\label{fig:naive-sketch}
\end{figure}

%bvand bvadd bvshl bvor bvxor bvlshr bvashr

\subsection{Generating Read-Write Sketches}\label{s:algorithm:rw}

The read-write sketch, $\RW(k, d, \iota, \arm_S, \arm_T, \Map)$, is based on the
observation that many practical source and target languages provide similar
functionality, so a source instruction $\iota$ can often be emulated with target
instructions that access the same parts of the state as $\iota$. For example,
the \cc{addi32} instruction from eBPF reads and writes only registers (not,
e.g., memory), and it can be emulated with RISC-V instructions that also touch
only registers (\autoref{s:overview}). Moreover, note that the semantics of
\cc{addi32} ignores the values of its $\mathid{src}$ and $\mathid{off}$
fields, and that the target RISC-V instructions do the same. Based on these
observations, our optimized sketch for \cc{addi32} would therefore consists of
instruction holes that allow only register-register instructions, with field
holes that exclude $\mathid{src}$ and $\mathid{off}$. We first formalize this
intuition with the notion of \emph{read and write sets}, and then describe how \jitsynth
applies such sets to create \RW sketches.\tighten

\paragraph{Read and write sets.} 
Read and write sets provide a compact way to summarize the semantics of an
abstract instruction $\iota$. This summary consists of a set of \emph{state
labels}, where a state label is one of $L_\regs$, $L_\mem$, and $L_\pc$
(\autoref{def:state-labels}). Each label in a summary set represents a state
component (registers, memory, or the program counter) that a concrete instance
of $\iota$ may read or write during some execution. We compute three such sets
of labels for every $\iota$: the read set $\Read{\iota}$, the write set
$\Write{\iota}$, and the write set $\Write{\iota, f}$ for each field $f$ of
$\iota$. \autoref{fig:rw-sets} shows these sets for the toy eBPF and RISC-V
instructions.\tighten

\begin{figure}\centering
  {\small
  \begin{tabular}{llll}
    \toprule
    $\iota$ & $\Read{\iota}$ & $\Write{\iota}$ & $\Write{\iota, \mathid{field}}$ \\
    \midrule
    \cc{addi32} & $\{L_\regs\}$ & $\{L_\regs\}$ & $\mathid{imm}$: $\{L_\regs\}$; $\mathid{off}$: $\emptyset$; $\mathid{src}$: $\emptyset$; $\mathid{dst}$: $\{L_\regs\}$ \\
    \cc{lui} & $\{L_\regs\}$ & $\{L_\regs\}$ & $\mathid{rd}$: $\{L_\regs\}$; $\mathid{imm20}$: $\{L_\regs\}$\\
    \cc{sb} & $\{L_\regs\}$ & $\{L_\mem\}$ & $\mathid{rs1}$: $\{L_\mem\}$; $\mathid{rs2}$: $\{L_\mem\}$; $\mathid{imm12}$: $\{L_\mem\}$ \\
    \bottomrule\end{tabular}
  }

\caption{Read and write sets for the \cc{addi32}, \cc{lui}, and \cc{sb} instructions from \autoref{fig:lang-instrs}.}\label{fig:rw-sets}
\end{figure}

The read set $\Read{\iota}$ specifies which components of the input state may
affect the execution of $\iota$ (\autoref{def:read-set}). For example, if
$\Read{\iota}$ includes $L_\regs$, then some concrete instance of $\iota$
produces different output states when executed on two input states that differ
only in register values. 
% More generally, $L_n \in \Read{\iota}$ means that the behavior of $\iota$ depends on the values read from the component $n$.\tighten
%
The write set $\Write{\iota}$ specifies which components of the output state may
be affected by executing $\iota$ (\autoref{def:write-set}). In particular, if
$\Write{\iota}$ includes $L_\regs$ (or $L_\mem$), then executing some concrete
instance of $\iota$ on an input state produces an output state with different
register (or memory) values. The inclusion of $L_\pc$ is based on a separate
condition, designed to distinguish jump instructions from fall-through
instructions. Both kinds of instructions change the program counter, but
fall-through instructions always change it in the same way. So,
$L_\pc\in\Write{\iota}$ if two instances of $\iota$ can write different values
to the program counter.
%
Finally, the field write set, $\Write{\iota, f}$, specifies the parts of the
output state are affected by the value of the field $f$; $L_n \in
\Write{\iota, f}$ means that two instances of $\iota$ that differ only in $f$
can produce different outputs when applied to the same input state.\tighten

\jitsynth computes all read and write sets from their definitions, by using the
host symbolic evaluator to reduce the reasoning about instruction semantics to
SMT queries. This reduction is possible because we assume that all ARM
interpreters are self-finitizing, as discussed in \autoref{s:overview}.

\begin{definition}[State Labels]\label{def:state-labels} 
  A \emph{state label} is an identifier $L_n$ where $n$ is a state component,
  i.e., $n\in \{\regs,\mem,\pc\}$. We write $N$ for the set of all state
  components, and $\mathcal{L}$ for the set of all state labels. We also use
  state labels to access the corresponding state components:
  $\Get{L_n}{\sigma} = n(\sigma)$ for all $n\in N$.
\end{definition}

\begin{definition}[Read Set]\label{def:read-set}
  Let $\iota\in\mathcal{I}$ be an abstract instruction in $(\mathcal{I},
  \Sigma, \mathcal{T}, \Phi)$. The \textup{read set} of $\iota$,
  $\Read{\iota}$, is the set of all state labels $L_n\in\mathcal{L}$ such that 
  $
    \exists p\in\prog{\iota}.\, 
    \exists L_w \in\Write{\iota}.\,
    \exists \sigma_a, \sigma_b \in\Sigma.\,
    (\Get{L_n}{\sigma_a} \neq \Get{L_n}{\sigma_b} \wedge 
    (\bigwedge_{m\in N \setminus \{ n \}} \Get{L_m}{\sigma_a} = \Get{L_m}{\sigma_b}) \wedge
    \Get{L_w}{\mathcal{T}(p,\sigma_a)} \neq \Get{L_w}{\mathcal{T}(p,\sigma_b)}.
  $\tighten
\end{definition}

\begin{definition}[Write Set]\label{def:write-set}
  Let $\iota\in\mathcal{I}$ be an abstract instruction in $(\mathcal{I}, \Sigma,
  \mathcal{T}, \Phi)$. The \textup{write set} of $\iota$, $\Write{\iota}$,
  includes the state label $L_n \in \{ L_\regs, L_\mem \}$ iff 
  $
    \exists p\in\prog{\iota}.\, 
    \exists \sigma \in\Sigma.\, 
    \Get{L_n}{\sigma} \neq \Get{L_n}{\mathcal{T}(p,\sigma)}
  $, 
  and it includes the state label $L_\pc$ iff 
  $
    \exists p_a, p_b\in\prog{\iota}.\, 
    \exists \sigma \in\Sigma.\,
    \Get{L_\pc}{\mathcal{T}(p_a,\sigma)} \neq \Get{L_\pc}{\mathcal{T}(p_b,\sigma)}.
  $\tighten
\end{definition}

\begin{definition}[Field Write Set]\label{def:field-set}
  Let $f$ be a field of an abstract instruction $\iota =
  (\mathid{op},\mathcal{F})$ in $(\mathcal{I}, \Sigma, \mathcal{T}, \Phi)$. The
  \textup{write set} of $\iota$ and $f$, $\Write{\iota, f}$, includes
  the state label $L_n \in \mathcal{L}$ iff 
  $
    \exists p_a, p_b\in\prog{\iota}.\, 
    \exists \sigma \in\Sigma.\ 
    (p_a.f \neq p_b.f) \wedge
    (\bigwedge_{g\in\mathid{dom}(\mathcal{F})\setminus\{f\}} p_a.g = p_b.g) \wedge
    \Get{L_n}{\mathcal{T}(p_a,\sigma)} \neq \Get{L_n}{\mathcal{T}(p_b,\sigma)}
  $, where $p.f$ denotes $F(f)$ for $p = (op, F)$. 
\end{definition}

% \begin{example}\label{ex:rw-sets}
%   \jitsynth computes the following read and write sets of $\iota = \cc{addi32}$:
%   $\Read{\iota} = \Write{\iota} = \Write{\iota, \mathid{dst}} = \Write{\iota,
%   \mathid{imm}} = \{L_\regs\}$, and $\Write{\iota, \mathid{src}} \\ = \Write{\iota,
%   \mathid{offset}} = \{\}$. The read and write sets for $\cc{lui}$ and its
%   fields are also $\{ L_\regs \}$, while for $\iota = \cc{sb}$, we get
%   $\Read{\iota}  = \{L_\regs \}$ and $\Write{\iota}  = \Write{\iota, f} =
%   \{L_\mem \}$ for each field $f$ of \cc{sb}.\tighten
% \end{example}

\paragraph{Using read and write sets.}

Given the read and write sets for a source instruction $\iota$ and target
instructions $\mathcal{I}_T$, \jitsynth generates the \RW sketch of length $k$
and depth $d$ by modifying the \Naive algorithm (\autoref{fig:naive-sketch}) as
follows. First, it restricts each target instruction hole (line
\ref{li:instruction-hole}) to choose an instruction $\iota_T\in\mathcal{I}_T$
with the same read and write sets as $\iota$, i.e., $\Read{\iota} =
\Read{\iota_T}$ and  $\Write{\iota} = \Write{\iota_T}$. Second, it restricts the
target field holes (line \ref{li:field-holes}) to use the source fields with the
matching field write set, i.e., the hole for a target field $f_T$ uses the
source field $f$ when $\Write{\iota_T, f_t} = \Write{\iota, f}$. For example,
given the sets from \autoref{fig:rw-sets}, the \RW instruction holes for
\cc{addi32} exclude \cc{sb} but include \cc{lui}, and the field holes for
\cc{lui} use only the $\mathid{dst}$ and $\mathid{imm}$ source fields. More
generally, the \RW sketch for \cc{addi32} consists of register-register
instructions over $\mathid{dst}$ and $\mathid{imm}$, as intended. This sketch
includes $2^{290}$ programs of length $k=5$ and depth $d\leq 3$, resulting in a
$2^{60}$ fold reduction in the size of the search space compared to the \Naive
sketch of the same length and depth.

% These restrictions capture the hypothesis that $\iota$ can be compiled to
% $\mathcal{I}_T$ instructions that manipulate state in the same way as $\iota$.
% This intuition, in turn, assumes that the states of the source and target
% machines are related by a state mapping (\autoref{def:state-equiv}) rather than
% an arbitrary equivalence relation.

\subsection{Generating Pre-Load Sketches}\label{s:algorithm:pld}

The pre-load sketch, $\LCS(k, d, \iota, \arm_S, \arm_T, \Map)$, is based on the
observation that hand-written JITs use macros or subroutines to generate
frequently used target instruction sequences. For example, compiling a source
instruction with immediate fields often involves loading the immediates into
scratch registers, and hand-written JITs include a subroutine that generates the
target instructions for performing these loads. The pre-load sketch shown in
\autoref{fig:lcs-sketch} mimics this structure.\tighten

In particular, \LCS generates a sequence of $m$ concrete instructions that load
the (used) immediate fields of $\iota$, followed by a sequence of $k-m$
instruction holes. The instruction holes can refer to both the source registers
(if any) and the scratch registers (via the arbitrary bitvector constants
included in the \emph{Field} holes). The function
$\mathid{Load}(\mathid{Expr}(p.f), \arm_T, \Map)$ returns a sequence of target
instructions that load the immediate $p.f$ into an unused scratch register. This
function itself is synthesized by \jitsynth using a variant of the \RW
sketch.\tighten

As an example, the pre-load sketch for \cc{addi32} consists of two $\mathid{Load}$
instructions (\cc{lui} and \cc{addiw} in the generated C code) and $k-2$
instruction holes.  The holes choose among register-register instructions in toy
RISC-V, and they can refer to the $\mathid{dst}$ register of \cc{addi32}, as
well as any scratch register. The resulting sketch includes $2^{100}$ programs
of length $k = 5$ and depth $d\leq 3$, providing a $2^{190}$ fold reduction in
the size of the search space compared to the \RW sketch.\tighten

\begin{figure}[h]
  \begin{algorithmic}[1] 
    \Function{PLD}
      {$k, d, \iota, \arm_S, \arm_T, \Map$}\Comment{$\iota\in\mathcal{I}_S, \arm_S = (\mathcal{I}_S, \ldots)$}
      \State $(\mathid{op}, \mathcal{F}) \gets \iota$, $(\mathid{I}_T, \ldots) \gets \arm_T$ \Comment{Source instruction, target instructions.}
      \State $p \gets \mathid{FreshId}()$ \Comment{Identifier for the compiler's input source instruction.}
      \State $\mathid{body}\gets []$  \Comment{The body of the compiler is a sequence with 2 parts:}
      \State $\mathid{imm} \gets \{ f \,|\, \mathcal{F}(f) = \BV{k} \text{ and } \Write{\iota,f} \neq \emptyset \}$ \Comment{(1) Load each relevant}
      \For{$f \in \mathid{imm}$} \Comment{source immediate into a free scratch register}
        \State $\mathid{body}\gets\mathid{body}\cdot \mathid{Load}(\mathid{Expr}(p.f), \arm_T, \Map)$ \Comment{using the load pseudoinstruction.}
      \EndFor
      % \If{$L_\pc\in\Read{\iota}$} \Comment{If $\iota$ reads the PC, load it into a free scratch register}
      %   \State $\mathid{body}\gets\mathid{body}\cdot \mathid{LoadPC}(\arm_T, \Map)$ \Comment{using the load PC pseudoinstruction.}
      % \EndIf
      \State $m \gets |\mathid{body}|$ \Comment{Let $m$ be the length of the load sequence.}
      \If{$m\geq k$ or $m = 0$} \Return $\bot$ \Comment{Return the empty sketch if $m\not\in (0..k)$.}
      \EndIf
      \For{$m \leq i < k$} \Comment{(2) Create $k-m$ target instruction holes, where the set}
        \State $I \gets \{\}$  \Comment{$I$ of choices for a target instruction hole includes}
        \For{$\iota_T\in\mathcal{I}_T, \iota_T = (\mathid{op}_T, \mathcal{F}_T)$} \Comment{all instructions from $\mathcal{I}_T$ that read-write}  
         \State $\mathid{rw}_T \gets  \Read{\iota_T} \times \Write{\iota_T}$ \Comment{the same state as $\iota$ or just registers.}
          \If{$\mathid{rw}_T = \Read{\iota} \times \Write{\iota}$ or $\mathid{rw}_T \subseteq \{L_\regs\} \times \{L_\regs\}$ }
            \State $\mathid{regs} \gets \{ f \,|\, \mathcal{F}(f) = \Reg \text{ and } \Write{\iota,f} \neq \emptyset \}$ \Comment{Any relevant}
            \State $E \gets \{\mathid{Expr}(p.f, \Map) \,|\, f\in\mathid{regs}\}$ \Comment{source register can appear in}
            \State $H \gets \{ f\mapsto \mathid{Field}(\mathcal{F}_T(f),d,E) \,|\, f\in\mathid{dom}(\mathcal{F}_T)\} $ \Comment{a target field hole,} 
            \State $I\gets I \cup \{ \mathit{Expr}((\mathid{op}_T, H),\Map) \}$ \Comment{and any constant register or value.}
          \EndIf
        \EndFor
        \State $\mathid{body} \gets \mathid{body} \cdot [\mathid{Choose}(I)]$ \Comment{Append a hole over $I$ to the body.}
      \EndFor
      % \If{$L_\pc\in\Write{\iota}$} \Comment{(3) If $\iota$ writes the PC, move a scratch register value}
      %   \State $\mathid{body}\gets\mathid{body}\cdot \mathid{StorePC}(\arm_T, \Map)$ \Comment{into the PC using a pseudoinstruction.}
      % \EndIf
      \State \Return $\mathid{Expr}((\lambda p\in\prog{\iota}\, .\, \mathid{body}),\Map)$ \Comment{A \minicompiler sketch for $\iota$.}
    \EndFunction
  \end{algorithmic}
  \caption{Pre-load sketch of length $k$ and maximum depth $d$ for $\iota$,
  $\arm_S$, $\arm_T$, and \Map. The $\mathid{Load}(E, \arm_T, \Map)$ function
  returns a sequence of target instructions that load the immediate value
  described by the expression $E$ into an unused scratch register; see
  \autoref{fig:naive-sketch} for descriptions of other helper
  functions.\tighten}\label{fig:lcs-sketch}
\end{figure}

\subsection{Solving Compiler Metasketches}\label{s:algorithm:solving}

\jitsynth solves the metasketch $\CMS(\iota, \arm_S, \arm_T, \Map)$ by applying
the host synthesizer to each of the generated sketches in turn until a
\minicompiler is found. If no \minicompiler exists in the search space, this
synthesis process runs forever. To check if a sketch $\mathcal{S}$ contains a
\minicompiler, \jitsynth would ideally ask the host synthesizer to solve the
following query, derived from Definitions \ref{def:mini-compiler}--\ref{def:state-equiv}: 
{\small
\[
    \exists C\in\mathcal{S} . \ 
    \forall \sigma_S\in\Sigma_S,\ \sigma_T\in\Sigma_T,\ p\in\prog{\iota}. \\ 
      \sigma_S \cong_{\Map} \sigma_T \Rightarrow
      \arm_S(p, \sigma_S) \cong_{\Map} \arm_T(C(p), \sigma_T)
\]}%
But recall that the state equivalence check $\cong_{\Map}$ involves universally
quantified formulas over memory addresses and register names. In principle,
these innermost quantifiers are not problematic because they range over finite
domains (bitvectors) so the formula remains decidable. In practice, however,
they lead to intractable SMT queries. We therefore solve a stronger soundness
query (\autoref{def:strong-mini-compiler}) that pulls these quantifiers out to
obtain the standard $\exists\forall$ formula with a quantifier-free body. The  
resulting formula can be solved with CEGIS~\cite{solar-lezama:sketch}, without 
requiring the underlying SMT solver to reason about quantifiers.

\begin{definition}[Strongly Sound \MiniCompiler]\label{def:strong-mini-compiler}
  Let $\arm_S = (\mathcal{I}_S, \Sigma_S, \mathcal{T}_S, \Phi_S)$ and $\arm_T =
  (\mathcal{I}_T, \Sigma_T, \mathcal{T}_T, \Phi_T)$ be two abstract register
  machines, $\cong_{\Map}$ an injective state equivalence relation on their
  states $\Sigma_S$ and $\Sigma_T$, and $C: \prog{\iota} \rightarrow
  \mathid{List}(\prog{\mathcal{I}_T})$ a function for some
  $\iota\in\mathcal{I}_S$. We say that $C$ is a \textup{strongly sound
  \minicompiler} for $\iota_{\Map}$ with respect to $\cong$ iff
  \begin{align*}\small
    & \forall \sigma_S\in\Sigma_S,\ \sigma_T\in\Sigma_T,\ p\in\prog{\iota}, \
    a\in\mathid{dom}(\mem(\sigma_S)),\ r\in\mathid{dom}(\regs(\sigma_S)). \\ 
    & \sigma_S \cong_{\Map,a,r} \sigma_T \Rightarrow
       \arm_S(p, \sigma_S) \cong_{\Map,a,r} \arm_T(C(p), \sigma_T)
\end{align*}
where $\cong_{\Map,a,r}$ stands for the $\cong_{\Map}$ formula with $a$ and $r$
as free variables. 
  \end{definition}

The \jitsynth synthesis procedure is sound and complete with respect to this
stronger query (\autoref{thm:synthesis-soundness-and-completeness}). The proof
follows from the soundness and completeness of the host synthesizer, and the
construction of the compiler metasketch. We discharge this proof using Lean theorem prover~\cite{moura:lean}.

\begin{theorem}[Strong soundness and completeness of \jitsynth]\label{thm:synthesis-soundness-and-completeness}
  Let $\mathcal{C} = \CMS(\iota, \arm_S, \arm_T, \Map)$ be the compiler
  metasketch for the abstract instruction $\iota$, machines $\arm_S$ and
  $\arm_T$, and the state mapping $\Map$. If \jitsynth terminates and returns a
  program $C$ when applied to $\mathcal{C}$, then $C$ is a strongly sound
  \minicompiler for $\iota$ and $\arm_T$ (soundness). If there is a strongly
  sound \minicompiler in the most general search space $\{\Naive(k, d, \iota,
  \arm_S, \arm_T, \Map) \,|\, k, d\in \mathbb{N}\}$, then \jitsynth will terminate on
  $\mathcal{C}$ and produce a program (completeness). 
\end{theorem}