### 2023-05-05 thesis.tex

\documentclass[12pt]{book}

%% Added by Jacob
\usepackage{makecell}
%\usepackage{amsthm}
%\theoremstyle{acmdefinition}
%\@ifundefined{example}{%
%}{}
%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\def\chapterautorefname{Chapter}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}

\usepackage{algorithm}
\usepackage{algpseudocode}
\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\algnewcommand\Not{\textbf{not} }
\newcommand{\algorithmautorefname}{Algorithm} % autoref

\usepackage{adjustbox}
\usepackage[headings]{fullpage}
\usepackage{xargs}
\usepackage{xspace}
\usepackage[doublespacing]{setspace}

\usepackage{microtype}
\usepackage[T1]{fontenc}

\usepackage{wrapfig}
\usepackage[font={sf}]{caption}
\usepackage[subrefformat=parens]{subcaption}

\usepackage{multirow}

\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{{_build/}{figures/}}

\usepackage{amsmath}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{example}{Example}[chapter]
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{proof}{Proof}[chapter]

\setlength{\marginparwidth}{2cm}
\usepackage[textsize=tiny]{todonotes}
%\renewcommand\max[2][]{\todo[color=orange, #1]{\sffamily #2}}

\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage[scaled=.85]{DejaVuSansMono}

%\usepackage{listings}
\usepackage[procnames]{listings}          % format code
\lstset{
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  stringstyle=\slshape\color{green!40!black},
  basicstyle=\ttfamily\small,
  language=Python,
  morekeywords={*, self},
  commentstyle=\slshape\color{black!60},
  tabsize=2,
}

\lstdefinelanguage{Rust}{
  sensitive,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  moredelim=[s][{\itshape\color[rgb]{0,0,0.75}}]{\#[}{]},
  morestring=[b]{"},
  alsodigit={},
  alsoother={},
  alsoletter={!},
  % keywords
  otherkeywords={=>},
  morekeywords={break, continue, else, for, if, in, loop, match, return, while},
  morekeywords={as, const, let, move, mut, ref, static, unsafe},
  morekeywords={dyn, enum, fn, impl, Self, self, struct, trait, type, use, where},
  morekeywords={crate, extern, mod, pub, super},
  morekeywords={abstract, alignof, become, box, do, final, macro,
    offsetof, override, priv, proc, pure, sizeof, typeof, unsized, virtual, yield},
  % traits
  morekeywords=[2]{Send},
  % types
  morekeywords=[3]{bool, char, f32, f64, i8, i16, i32, i64, isize, str, u8, u16, u32, u64, unit, usize, i128, u128},
}%

\def\mytitle{Program Synthesis for Systems Developers}
\def\myauthor{Jacob Van Geffen}
\def\year{2023}

\title{\mytitle}
\author{\myauthor}

% TODO write thesis statement here
\newcommand\thesisstmt{%
  e-graphs and equality saturation are compelling techniques for
  program representation and manipulation that should now be considered
  for programming tools across many domains.
}

\newcommand\Thesisstmt{\expandafter\MakeUppercase \thesisstmt}

\input{macros}
\input{dmacros}
\input{data/nr}

\begin{document}
\pagestyle{empty}

% title and copyright pages
\begin{center}
  {\huge \mytitle}
  \vfill

  {\Large \myauthor}
  \vfill

  \begin{spacing}{1}
    A dissertation \\
    submitted in partial fulfillment of the \\
    requirements for the degree of
  \end{spacing}
  \vfill

  Doctor of Philosophy
  \vfill

  University of Washington \\
  \year
  \vfill

  Reading Commitee: \\
  Emina Torlak, Chair \\
  Xi Wang \\
  Zachary Tatlock \\
  Shane Steinert-Threlkeld
  \vfill

  \begin{spacing}{1}
    Program Authorized to Offer Degree: \\
    \cse
  \end{spacing}
  \clearpage

  \textcopyright{} Copyright \year\\
  \myauthor
  \clearpage
\end{center}

\pagestyle{plain}
\setcounter{page}{1}
\pagenumbering{roman}

% abstract page
\begin{center}
  University of Washington \\[1em]
  \textbf{Abstract}        \\[1em]
  \mytitle                 \\[1em]
  \myauthor                \\[1em]

  % Supervisory Committee: \\[-0.5em]
  % Luis Ceze, Chair \qquad
  % Adriana Schulz   \qquad
  % Zachary Tatlock  \\[-0.5em]
  Chair of the Supervisory Committee: \\[-0.5em]
  Emina Torlak \\[-0.5em]
  \pgas
  \\[2em]
\end{center}
\input{abstract}

\begin{spacing}{1.5}
  \tableofcontents
\end{spacing}

% TODO write acknowledgements HERE
\chapter*{Acknowledgments}

% Emina
Over the past five years, I've had the priviledge of working with
of Emina Torlak. She is everything I could have possibly asked for
in a mentor. From getting into the Rosette weeds to career chats over Molly Moon's,
she always seems to know exactly what I need to hear and when I need to hear it.
Thanks for always believing me and keeping me moving forward.

% James
Before I ever started my PhD at UW, I saw James Bornholt give a talk at PLDI 2017
about his work synthesizing memory models. Of course I loved the talk,
but at the time, I had no idea how much of an impact James' work would
have on my life. Without the groundwork of James' ridiculously prolific
career, this thesis would not exist. As if this wasn't enough, James has
also been an excellent mentor and friend throughout my time at UW. 
Thanks for working with me, James.

% Isil
My path to writing this thesis feels like a series of insanely lucky
happenstances. Starting my undergraduate degree at UT Austin
the same semester that Isil Dillig became a faculty member at UT wasn't the
first or last bit of this luck, but it certainly was one of the most impactful.
Isil is the reason that I started my program synthesis research journey, and I
likely never would have pursued a PhD without her guidance.
% TODO conclude?

% Collabs
I'm extremely grateful to everyone that I've worked with during my time at UW.
Without fail, Xi Wang always asks the most important and challenging questions,
and I wouldn't be half the researcher I am today without him.
Luke Nelson brought me up to speed when I started in the UNSAT group at UW,
and for that I'll always be thankful.
Additionally, I'd like to thank Zach Tatlock and Shane Steinert-Threlkeld
for agreeing to be part of my thesis committee.
Zach, thanks for making me feel at home in new, scary surroundings.

% Unsat + PLSE
I doubt I could have finished this PhD with my sanity intact without the support
of all my friends from the UNSAT group, PLSE lab, Systems lab, and Architecture
group. There are so many amazing people that have welcomed me into these communities,
but especially to
Oliver Flatt,
Ben Kushigian,
Katie Lim,
Ellis Michael,
Samantha Miller,
Chandrakana Nandi,
Pratyush Patel,
Dan Petrisko,
Sorawee Porncharoenwase,
Max Ruttenberg,
Gus Smith,
Ewin Tang,
and Max Willsey,
thank you for making these past five years a blast.

% Fam + Madeline
Finally, I cannot emphasize how crucial the love and support from my family
has been over the past five years. My parents Shiela and Steve and my brothers
Daniel and Sammy have been a non-stop source of encouragement.
Madeline, I'll never be able to thank you enough for sticking through this whole
journey with me, but I will certainly try. Thank you for always being on my team.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\pagestyle{headings}
\setcounter{page}{1}
\pagenumbering{arabic}
\renewcommand{\chaptermark}[1]{\markboth{\sc{\chaptername\ \thechapter.\ #1}}{}}
\renewcommand{\sectionmark}[1]{\markright{\sc{\thesection.\ #1}}{}}
% \listoftodos
\input{chapters/intro}
\input{chapters/jitsynth/jitsynth}
\input{chapters/depsynth/depsynth}

\clearpage
\addcontentsline{toc}{chapter}{Bibliography}
\singlespacing
\bibliographystyle{alpha}
\bibliography{references}

\end{document}