### 2023-05-10 chapters/depsynth/improved-search.tex

\section{Improving Happens-Before Graph Search}
\label{s:improved-search}
A major bottleneck in the time \depsynth takes to find a complete set of
dependency rules comes from the time it takes to search for a minimal
happens-before graph for a given litmus test.
This overhead becomes exacerbated for more complex storage systems
that require larger litmus tests to describe their behavior.
This section explores techniques for reducing this overhead.
% The

In order to minimize the happens-before graph search time,
we took two approaches:
\begin{enumerate}
  \item Encoding the problem in the theory of ordering consistency (\tord)\todo{cite}
        and search for happens-before graphs with an existing \tord solver
  \item Optimizing existing
\end{enumerate}
The next subsections describe these approaches and their respective effectiveness
in detail.