### 2023-05-11 chapters/concl.tex
Conclusion; make file

\chapter{Conclusion}
% TODO recap
This dissertation has presented two tools --- \jitsynth and \depsynth\ ---
along with new metaprogramming abstractions and system decompositions
that enable the tools to automate system design and implementation for their
respective domains.
\cref{c:jitsynth} demonstrated how \jitsynth frames the synthesis task
as a search for per-instruction compilers over \textit{abstract register machines},
taking advantage of new minicompiler metasketches to efficiently synthesize
real-world JITs.
In \cref{c:depsynth}, we introduced a new \textit{angelic} programming model for
crash consistency, along with a tool \depsynth that synthesizes
\textit{dependency rules} to enforce crash consistency under this model.
We also evaluated \depsynth on \shardstore, a production key-value storage system
at Amazon, to show the effectiveness of this technique.
Both works demonstrate the effectiveness of program synthesis in
designing correct and performant systems.

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\todo{cite crashmonkey and memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discussed existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Properties for staking protocols are specified in a fairly straightforward mannor,
but complex invariants on the blockchain state make implementing staking mechanisms difficult.
\todo{cite https://academic.oup.com/rfs/article-abstract/34/3/1156/5868423, https://link.springer.com/chapter/10.1007/978-3-319-67816-0_17, https://link.springer.com/chapter/10.1007/978-3-319-63688-7_12}
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels,\todo{cite}
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.