### 2023-05-15 figs/rules.tex
Split up big figure

\begin{algorithmic}[1]
\algrestore{depsynth}

\Function{\rulesfortest}{$\test = \langle P_\emph{initial}, P_\emph{main} \rangle$, \impl, \consist}
    \State $\wr \gets \{w\ | \ w \in \evaluate{P_\emph{main}}{\impl} \}$
    \State \Return $\Call{\phaseone}{\tests, [], \wr, \impl, \consist}$
\EndFunction

\vspace{0.8em}

\Function{\phaseone}{\test, \ord, \wr, \impl, \consist}  \Comment{Search for total orders over writes}  \label{fig:rules:phase1}
    \If{$\wr = \emptyset$}
        \State $\gr \gets \{ (\ord[i], \ord[j]) \ | \ 0 \leq i < j < |\ord| \}$
        \State \Return \Call{\phasetwo}{\tests, \gr, \impl, \consist}  \Comment{\gr is a total order; minimize it in Phase 2} \label{fig:rules:phase1tophase2}
    \EndIf
    \For{$w \in \wr$}
        \State $\ord' \gets \ord + [w]$
        \State $\wr' \gets \wr \setminus \{ w \}$
        \State $\gr \gets \{ (\ord[i], \ord[j]) \ | \ 0 \leq i < j < |\ord| \} \, \cup$ 
            \Statex $\hspace{6em}\{ (w_1, w_2) \ | \ w_1 \in \ord \land w_2 \in \wr \} \, \cup$ 
            \Statex $\hspace{6em}\{ (w_1, w_2) \ | \ w_1, w_2 \in \wr \}$  \label{fig:rules:phase1angelic}
         \If{$\neg \crashconsistentalg(\test, \Call{\rulesforgraph}{\gr}, \impl, \consist)$}  \label{fig:rules:phase1sufficient}
            \State \textbf{continue} \label{fig:rules:phase1continue}
        \EndIf
        \State $\ruleset \gets \Call{\phaseone}{\test, \ord', \wr', \impl, \consist}$
        \If{$\ruleset \neq \bot$}
            \State \Return \ruleset
        \EndIf
    \EndFor
    \State \Return $\bot$
\EndFunction

\vspace{0.8em}

\Function{\phasetwo}{\test, \gr, \impl, \consist}  \Comment{Minimize graph \gr by removing individual edges}  \label{fig:rules:phase2}
    \State $\ruleset \gets \Call{\rulesforgraph}{\gr}$
    \If{$\neg \Call{\crashconsistentalg}{\test, \ruleset, \impl, \consist}$}  \label{fig:rules:phase2sufficient}
        \State \Return $\bot$
    \EndIf
    \For{$(w_1, w_2) \in \gr$}  \Comment{Try removing each edge from \gr}  \label{fig:rules:phase2loop}
        \State $\gr' \gets \gr \setminus \{(w_1, w_2)\}$  \label{fig:rules:phase2remove}
        \State $\ruleset' \gets \Call{\phasetwo}{\test, \gr', \impl, \consist}$ \label{fig:rules:phase2recurse}
        \If{$\ruleset' \neq \bot$}
            \State \Return \ruleset'
        \EndIf
    \EndFor
    \If{$\Call{Acyclic}{\ruleset}$}
        \State \Return \ruleset \Comment{\gr makes \test crash consistent and no subgraph of \gr suffices}
    \Else
        \State \Return $\bot$
    \EndIf
\EndFunction

\vspace{0.8em}

\algstore*{depsynth}
\end{algorithmic}