### 2023-05-15 figs/rulesforgraph.tex

\begin{algorithmic}[1]
\algrestore{depsynth}

\Function{\rulesforgraph}{\gr}  \Comment{Generalize a happens-before graph into dependency rules}  \label{fig:rules:rulesforgraph}
    \State $\ruleset \gets \{\}$
    \For{$(w_1, w_2) \in \gr$}
        \State $\langle n_1, t_1 \rangle \gets \Call{Label}{w_1}$  \Comment{Get label $l_1 = \langle n_1, t_1 \rangle$ for write $w_1 = \texttt{write}(a_1, s_1, l_1)$}
        \State $\langle n_2, t_2 \rangle \gets \Call{Label}{w_2}$
        \If{$t_1 < t_2$}
            \State $\ruleset \gets \ruleset \cup \{ \deprule{n_2}{n_1}{>} \}$  \Comment{Invert the order, as a rule $\deprule{n_a}{n_b}{p}$ says $n_a$ happens \textbf{after} $n_b$}
        \ElsIf{$t_1 = t_2$}
            \State $\ruleset \gets \ruleset \cup \{ \deprule{n_2}{n_1}{=} \}$
        \Else
            \State $\ruleset \gets \ruleset \cup \{ \deprule{n_2}{n_1}{<} \}$
        \EndIf
    \EndFor
    \State \Return \ruleset
\EndFunction

\algstore*{depsynth}
\end{algorithmic}