### 2023-05-15 figs/phase2.tex
Increase font size for depsynth algorithms

\begin{algorithmic}[1]
\algrestore{depsynth}

\Function{\phasetwo}{\test, \gr, \impl, \consist}  \Comment{Minimize graph \gr by removing individual edges}  \label{fig:rules:phase2}
    \State $\ruleset \gets \Call{\rulesforgraph}{\gr}$
    \If{$\neg \Call{\crashconsistentalg}{\test, \ruleset, \impl, \consist}$}  \label{fig:rules:phase2sufficient}
        \State \Return $\bot$
    \EndIf
    \For{$(w_1, w_2) \in \gr$}  \Comment{Try removing each edge from \gr}  \label{fig:rules:phase2loop}
        \State $\gr' \gets \gr \setminus \{(w_1, w_2)\}$  \label{fig:rules:phase2remove}
        \State $\ruleset' \gets \Call{\phasetwo}{\test, \gr', \impl, \consist}$ \label{fig:rules:phase2recurse}
        \If{$\ruleset' \neq \bot$}
            \State \Return \ruleset'
        \EndIf
    \EndFor
    \If{$\Call{Acyclic}{\ruleset}$}
        \State \Return \ruleset \Comment{\gr makes \test crash consistent and no subgraph of \gr suffices}
    \Else
        \State \Return $\bot$
    \EndIf
\EndFunction

\algstore*{depsynth}
\end{algorithmic}